\simname 2.0 User's Guide

### 2011-10-22 doc/manual.tex
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/trunk@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/gems_interface@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515
update license text for booksim
git-svn-id: https://nocs.stanford.edu/cgi-bin/svn.cgi/booksim2.0/branches/pbooksim@4080 ce7aba0d-9f81-40a6-a495-cecae84b0515

% Copyright (c) 2007-2009, Trustees of The Leland Stanford Junior University
% All rights reserved.
% 
% Redistribution and use in source and binary forms, with or without 
% modification, are permitted provided that the following conditions are met:
% 
% Redistributions of source code must retain the above copyright notice, this 
% list of conditions and the following disclaimer.
% Redistributions in binary form must reproduce the above copyright notice, 
% this list of conditions and the following disclaimer in the documentation 
% and/or other materials provided with the distribution.
% Neither the name of the Stanford University nor the names of its contributors
% may be used to endorse or promote products derived from this software without
% specific prior written permission.
% 
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS" 
% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE 
% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE 
% ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE 
% LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR 
% CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF 
% SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS 
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN 
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE) 
% ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE 
% POSSIBILITY OF SUCH DAMAGE.

@@

\author{Nan Jiang, George Michelogiannakis, Daniel Becker, Brian Towles and William J. Dally}
\title{\simname 2.0 User's Guide}

@@

The latest official release of the simulator can be checked out from Booksim's sourceforge subversion (SVN) repository. Make sure SVN is installed on your machine and run the following command:
\begin{verbatim}
svn co https://booksim.svn.sourceforge.net/svnroot/booksim
\end{verbatim}

The simulator files should now be in the \texttt{booksim/trunk/src/} directory.
The simulator itself is written in C++ and has been specifically tested with GNU's G++ compiler (version $\ge3$).
The front end of the simulator uses LEX and YACC generated parser to process the simulator configuration file; however, unless you plan on making changes to the front end parser, LEX and YACC are not needed.
The \texttt{Makefile} should be edited so that the first few lines reflect the correct paths to the tools for your particular system.
The default \texttt{Makefile} should work on the Stanford Leland machines.
Type \texttt{make} to build the simulator.

A note for Windows users:
The above instructions have been tested to work with the newest version of Cygwin.

@@

The simulator is invoked using the following command line:
\begin{verbatim}
  ./booksim [configfile]
\end{verbatim}

The parameter \texttt{configfile} is a file that contains configuration information for the simulator.
So, for example, to simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube) network on uniform traffic, a configuration such as the one shown in Figure~\ref{fig:config_example} could be used.
This particular example configuration file can be found in the \texttt{examples/torus88} directory.

@@

In addition to specifying the topology, the configuration file also contains basic information about the routing algorithm, flow control, and traffic.
This simple example uses dimension-order routing and four virtual channels.
The \texttt{injection\_rate} parameter is added to tell the simulator to inject (on average) 0.15 packets per simulation cycle per node.
Packet size defaults to a single flit.
Also, any line of the configuration file that begins with \texttt{//} is treated as a comment and ignored by the simulator.
A detailed list of configuration parameters is given in Section~\ref{sec:config_params}.
Any parameters not specified by the user will take on default values.
The default value for every parameter in the simulator is specified in the file \texttt{booksim\_config.cpp}.

@@

\begin{figure}
\begin{verbatim}
BEGIN Configuration File
Name: examples/torus88
// Topology
topology = torus;
k = 8;
n = 2;
// Routing
routing_function = dim_order;
// Flow control
num_vcs = 4;
// Traffic
traffic = uniform;
injection_rate = 0.15;

END Configuration File
0%=================================
% Average latency = 10.3206
% Accepted packets = 0.113 at node 38 (avg = 0.149031)
lat(1) = 10.3206;
bins = [ 0 ...999];
freq = [ 0 0 145 4 550 52 988 137 1439 285 1656 428 1359 417 958 294 453 163 143
 49 10 7 1 0 ....

....

@@

....

% Draining all recorded packets ...
% Draining remaining packets ...
Time taken is 6030 cycles
====== Traffic class 0 ======
Overall average latency = 10.2897 (1 samples)
Overall average accepted rate = 0.14834 (1 samples)
Overall min accepted rate = 0.129043 (1 samples)
Average hops = 4.99839 (57153 samples)
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

@@

\section{Example}
\label{sec:examples}

One of the most basic performance measures of any interconnection
network is its latency versus offered load.
Figure~\ref{fig:lat_vs_load} shows a simple configuration file for
making this measurement in a 8-ary 2-mesh network under the transpose
traffic pattern.  This configuration was used to generate a single data point in Figure 25.2
in PPIN.  The particular configuration accounts for some small delays
and pipelining of the input-queued router and also introduces a small
input speedup to account for any inefficiencies in allocation.  By
running simulations for many increments of \texttt{injection\_rate},
the average latency curve can be found.  Then, to compare the
performance of dimension-order routing against several other routing
algorithms, for example, the \texttt{routing\_function} option can be
changed.

@@

sim_type       = latency;
injection_rate = 0.01;
\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

@@

\item[flattened butterfly] A topology based on the paper "Flattened butterfly: a cost-efficient topology for high-radix networks" ISCA 2007

@@

\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung from  EE382C several years ago.

@@

\subsubsection{Injection mode}
The rate at which packets are injected into the simulator is set using
the \texttt{injection\_rate} option.  The simulator's cycle time is a
flit cycle, the time it takes a single flit to be injected at a
source, and the injection rate is specified in packets per flit cycle.
For example, setting \texttt{injection\_rate = 0.25} means that each
source injects a new packet in one out of every four simulator cycles.
The unit of \texttt{injection\_rate} can optionally be changed to flits per cycle by 
setting \texttt{injection\_rate\_uses\_flits} to 1.

The injection process can further be specified as either Bernoulli process 
(\texttt{injection\_process = bernoulli}) or an on-off process
(\texttt{injection\_process = on\_off}).  The burstiness of the latter
is controlled via the \texttt{burst\_alpha} and
\texttt{burst\_beta} parameters.  See PPIN Section 24.2.2 for a
description of the on-off process and its parameters.

\subsubsection{Request-reply traffic}
By default, all packets that are injected into the network have the same, 
fixed length. The number of flits per packet is set using the
\texttt{const\_flits\_per\_packet} option.

Alternatively, the simulator can be configured to generate request-reply 
traffic. In this mode, the traffic manager injects read and write requests 
into the network; when a destination node receives such a request 
packet, it returns a reply packet of the same type. Injection of reply 
packets has priority over injection of new requests packets. 
Consequently, the effective overall network load is generated by both 
the injected requests and the automatically generated replies. 
Request-reply traffic ignores the \texttt{const\_flits\_per\_packet} 
option; instead, packet sizes are determined by the 
\texttt{\{read|write\}\_\{request|reply\}\_size} options. Furthermore, 
the mapping of packet types to VCs can be customized using the 
\texttt{\{read|write\}\_\{request|reply\}\_\{begin|end\}\_vc} options.

\subsubsection{Traffic patterns}
The simulator also supports several different traffic patterns that
are specified using the \texttt{traffic} option.  To describe these
patterns, we use the same notation of PPIN Section 3.2: $s_i$ ($d_i$)
denotes the $i^\textrm{th}$ bit of the source (destination) address
whereas $s_x$ ($d_x$) denotes the $x^\textrm{th}$ radix-$k$ digit of
the source (destination) address.  The bit length of an address is $b
= \log_2 N$, where $N$ is the number of nodes in the network.

\begin{opt_list}{trafficpatterns}
\item[uniform] Each source sends an equal amount of traffic to each
destination (\texttt{traffic = uniform}).
\item[bitcomp] Bit complement. $d_i = \neg s_i$.
\item[bitrev] Bit reverse. $d_i = s_{b-i-1}$.
\item[shuffle] $d_i = s_{i-1 \mod b}$.
\item[transpose] $d_i = s_{i+b/2 \mod b}$.
\item[tornado] $d_x = s_x + \lceil k/2 \rceil - 1 \mod k$.
\item[neighbor] $d_x = s_x + 1 \mod k$.
\item[randperm] Random permutation.  A fixed permutation traffic
pattern is chosen uniformly at random from the set of all
permutations.  The seed used to generate this permutation is set by
the \texttt{perm\_seed} option.  So, randomly selecting values for
\texttt{perm\_seed} gives a random sampling of permutations while a
fixed value of \texttt{perm\_seed} allows the same permutation to be
used for several experiments.
\end{opt_list}

@@

\appendix
\section{Random number generation}

>>>>

% Copyright (c) 2007-2011, Trustees of The Leland Stanford Junior University
% All rights reserved.
%
% Redistribution and use in source and binary forms, with or without
% modification, are permitted provided that the following conditions are met:
%
% Redistributions of source code must retain the above copyright notice, this
% list of conditions and the following disclaimer.
% Redistributions in binary form must reproduce the above copyright notice,
% this list of conditions and the following disclaimer in the documentation
% and/or other materials provided with the distribution.
%
% THIS SOFTWARE IS PROVIDED BY THE COPYRIGHT HOLDERS AND CONTRIBUTORS "AS IS"
% AND ANY EXPRESS OR IMPLIED WARRANTIES, INCLUDING, BUT NOT LIMITED TO, THE
% IMPLIED WARRANTIES OF MERCHANTABILITY AND FITNESS FOR A PARTICULAR PURPOSE
% ARE DISCLAIMED. IN NO EVENT SHALL THE COPYRIGHT OWNER OR CONTRIBUTORS BE
% LIABLE FOR ANY DIRECT, INDIRECT, INCIDENTAL, SPECIAL, EXEMPLARY, OR
% CONSEQUENTIAL DAMAGES (INCLUDING, BUT NOT LIMITED TO, PROCUREMENT OF
% SUBSTITUTE GOODS OR SERVICES; LOSS OF USE, DATA, OR PROFITS; OR BUSINESS
% INTERRUPTION) HOWEVER CAUSED AND ON ANY THEORY OF LIABILITY, WHETHER IN
% CONTRACT, STRICT LIABILITY, OR TORT (INCLUDING NEGLIGENCE OR OTHERWISE)
% ARISING IN ANY WAY OUT OF THE USE OF THIS SOFTWARE, EVEN IF ADVISED OF THE
% POSSIBILITY OF SUCH DAMAGE.

@@

\author{Ted Jiang, George Michelogiannakis, Daniel Becker, Brian Towles and William J. Dally}
\title{\simname 2.0 User's Guide}

@@

The latest version of the simulator is available from
\texttt{http://cva.stanford.edu} as a compressed tar archive.  UNIX/Linux
users can extract this archive using the tar utility
\begin{verbatim}

  tar xvfz booksim-2.0.tar.gz  
\end{verbatim}

Windows users can use a compression program such as Winrar to extract
the archive.

The simulator itself is written in C++ and has been specifically
tested with GNU's G++ compiler (version $\ge3$).  In addition, both a
LEX and YACC tool (also known as FLEX and BISON) are needed to create
the configuration parser.  These are standard tools in any UNIX/Linux
development environment.  It is suggested that Windows users download
the CYGWIN versions (\texttt{http://www.cygwin.com}) of these UNIX
development tools to simplify their compilation process.  The
\texttt{Makefile} should be edited so that the first lines give the
paths to the tools.  At Stanford, for example, the compiler, YACC, and
LEX are stored in the \texttt{/usr/pubsw/bin} directory.  The
\texttt{Makefile} reflects this:
\begin{verbatim}
CPP    = /usr/pubsw/bin/g++
YACC   = /usr/pubsw/bin/bison -d
LEX    = /usr/pubsw/bin/flex
\end{verbatim}
Then, the simulator can be compiled by running \texttt{make} in the
directory that contains the \texttt{Makefile}.

@@

The syntax of the simulator is simply
\begin{verbatim}
  ./booksim [configfile]
\end{verbatim}
The optional parameter \texttt{configfile} is a file that contains
configuration information for the simulator.  So, for example, to
simulate the performance of a simple $8 \times 8$ torus (8-ary 2-cube)
network on uniform traffic, a configuration such as the one shown in
Figure~\ref{fig:config_example} could be used. This particular configuration is stored in examples/torus88.

@@

In addition to specifying the topology, the configuration file also
contains basic information about the routing algorithm, flow control,
and traffic.  This simple example uses dimension-order routing and and four
virtual channels.  The \texttt{injection\_rate} parameter
is added to tell the simulator to inject 0.15 packets per simulation
cycle per node.  The packets are default to size of 1 flit.  Also, any line of the configuration
that begins with \texttt{//} is treated as a comment and ignored by
the simulator.  A detailed list of configuration parameters is given in
Section~\ref{sec:config_params}.

@@

\begin{figure}
\begin{verbatim}
% Average latency = 6.02008
% Accepted packets = 0.11 at node 52 (avg = 0.147094)
% latency change    = 1
% throughput change = 1
>>>>>>> variant B
BEGIN Configuration File
Name: examples/torus88
// Topology
topology = torus;
k = 8;
n = 2;
// Routing
routing_function = dim_order;
// Flow control
num_vcs = 4;
// Traffic
traffic = uniform;
injection_rate = 0.15;

END Configuration File
Unable to open flit watch file, continuing with simulation
0%=================================
% Average latency = 10.6252
% Accepted packets = 0.119 at node 5 (avg = 0.148875)
lat(1) = 10.6252;
bins = [ 0 ...999];
freq = [ 0 0 124 16 497 71 920 166 1336 303 1553 436 1351 
406 926 328 537 213 173 77 43 24 13 6 5 2 1 1 0 ...]

...

@@

...

% Draining all recorded packets ...
% Draining remaining packets ...
Time taken is 6031 cycles
====== Traffic class 0 ======
Overall average latency = 10.658 (1 samples)
Overall average accepted rate = 0.148383 (1 samples)
Overall min accepted rate = 0.134279 (1 samples)
Average hops = 5.01172 (57148 samples)
\end{verbatim}
\caption{Simulator output from running the \texttt{examples/torus88}
configuration file.}
\label{fig:sim_output}
\end{figure}

@@

\section{Examples}
\label{sec:examples}

One of the most basic performance measures of any interconnection
network is its latency versus offered load.
Figure~\ref{fig:lat_vs_load} shows a simple configuration file for
making this measurement in a 8-ary 2-mesh network under the transpose
traffic pattern.  This configuration was used to generate Figure 25.2
in PPIN.  The particular configuration accounts for some small delays
and pipelining of the input-queued router and also introduces a small
input speedup to account for any inefficiencies in allocation.  By
running simulations for many increments of \texttt{injection\_rate},
the average latency curve can be found.  Then, to compare the
performance of dimension-order routing against several other routing
algorithms, for example, the \texttt{routing\_function} option can be
changed.

@@

sim_type       = latency;
injection_rate = 0.005;
\end{verbatim}
\caption{A typical configuration file (\texttt{examples/mesh88\_lat})
for creating a latency versus offered load curve for a 8-ary 2-mesh
network.}
\label{fig:lat_vs_load}
\end{figure}

@@

\item[flattened butterfly] A topology based on ***

@@

\item[cmo] Concentrated Multi-dimensional Octagon created by Sang Kyun Kim and Wook-Jin Chung.

@@

The two traffic modes, injection rate or batch, can be set by the integer paramter \texttt[use\_read\_write], where 0 means injection mode. Injection mode is on by default.

\subsubsection{Injection mode}
The rate at which packkets are injected into the simulator is set using
the \texttt{injection\_rate} option.  The simulator's cycle time is a
flit cycle, the time it takes a single flit to be injected at a
source, and the injection rate is specified in packets per flit cycle.
For example, setting \texttt{injection\_rate = 0.25} means that each
source injects a new packet one of every four simulator cycles.  The
injection process can also be specified as either Bernoulli
(\texttt{injection\_process = bernoulli}) or an on-off process
(\texttt{injection\_process = on\_off}).  The burstiness of the latter
injection process is controlled via the \texttt{burst\_alpha} and
\texttt{burst\_beta} parameter.  See PPIN Section 24.2.2 for a
description of the on-off process and its parameters.

The unit of injection is packets, which may be comprised of many

flits. All packets in injection mode are identical. The number of flits per packet is set using the
\texttt{const\_flits\_per\_packet} option.

\subsubsection[Batch mode]
Batch mode is enabled by setting \texttt{use\_read\_write= 1}. In batch mode, \texttt[injection\_rate] paramter is ignored. Instead, traffic are injected into the network as fast as possible until \texttt{batch\_size} of request packets are generated by each node. When a destination node receives a request packet, it returns a reply packet. The total number of packets injected in batch mode is 2$\times$ \texttt{batch\_size} per node. The parameter \texttt[max\_outstanding] specifies the maximum number of (request or reply) packets waiting at a node to be injected into the network.

\subsubsection[Traffic Patterns]
The simulator also supports several different traffic patterns that
are specified using the \texttt{traffic} option.  To describe these
patterns, we use the same notation of PPIN Section 3.2: $s_i$ ($d_i$)
denotes the $i^\textrm{th}$ bit of the source (destination) address
whereas $s_x$ ($d_x$) denotes the $x^\textrm{th}$ radix-$k$ digit of
the source (destination) address.  The bit length of an address is $b
= \log_2 N$, where $N$ is the number of nodes in the network.

\begin{opt_list}{trafficpatterns}
\item[uniform] Each source sends an equal amount of traffic to each
destination (\texttt{traffic = uniform}).
\item[bitcomp] Bit complement. $d_i = \neg s_i$.
\item[bitrev] Bit reverse. $d_i = s_{b-i-1}$.
\item[shuffle] $d_i = s_{i-1 \mod b}$.
\item[transpose] $d_i = s_{i+b/2 \mod b}$.
\item[tornado] $d_x = s_x + \lceil k/2 \rceil - 1 \mod k$.
\item[neighbor] $d_x = s_x + 1 \mod k$.
\item[randperm] Random permutation.  A fixed permutation traffic
pattern is chosen uniformly at random from the set of all
permutations.  The seed used to generate this permutation is set by
the \texttt{perm\_seed} option.  So, randomly selecting values for
\texttt{perm\_seed} gives a random sampling of permutation while a
fixed value of \texttt{perm\_seed} allows the same permutation to be
used for several experiments.
\end{opt_list}

@@

\appendix

\section{Random number generation}
